---
title: "ESM 244 Assignment 2"
author: "Danny Ettelson"
date: "2/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r loadlib, message=FALSE, warning=FALSE}

library(tidyverse)
library(stargazer)
library(lubridate)
library(boot)
library(knitr)
cetaceans <- read_csv("captive_cetacean.csv")

```

### Task 1: Data wrangling and visualization - cetaceans in captivity



```{r captivity, fig.align='center', fig.cap="\\label{fig:figs}Number of cetaceans born in captivity 1950 - 2018. Births where artificial insemination was used are shown in blue", message=FALSE, warning=FALSE}


cetaceans$birthYear <- as.numeric(cetaceans$birthYear)


cetaceans_fathers <- cetaceans %>% 
  drop_na(father) %>% 
  drop_na(birthYear) %>% 
  group_by(father) %>% 
  mutate(AI = grepl("A.I.", father))


ggplot(cetaceans_fathers) +
  geom_histogram(aes(x = birthYear, fill = AI), binwidth = 1) +
  labs(x = "Birth Year", y = "Count")+
  ggtitle("Cetacean Births in Captivity 1950 - 2018") +
  #scale_fill_manual(labels = c("True", "False"),values = c("blue", "red")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5),legend.position="bottom") +
  guides(fill=guide_legend(title="Use of Artificial Insemination")) 
  #theme_classic()





```

```{r releases,fig.align='center', fig.cap="\\label{fig:figs}Release of cetaceans from captivity colored by species.", message=FALSE, warning=FALSE}

cetaceans$statusDate <- mdy(cetaceans$statusDate)

cetaceans_release <- cetaceans %>% 
  filter(status == "Released") %>% 
  drop_na(statusDate) %>% 
  mutate(year_released = year(statusDate))

ggplot(cetaceans_release) +
  geom_histogram(aes(x = year_released, fill = species), binwidth = 1) +
  scale_fill_brewer(palette = "Set2") +  
  labs(x = "Year Released", y = "Count") +
  guides(fill=guide_legend(title="Species")) +
  ggtitle("Cetacean Releases from Captivity") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5),legend.position="bottom")



```

### Task 2: Parameter Estimation – Wild Fish Catch


```{r message=FALSE, warning=FALSE, fig.align='center'}
fish_catch <- read_csv("fish_catch.csv")

fish_catch <- drop_na(fish_catch)
colnames(fish_catch) <- c("year", "wild", "farmed", "total")
fish_catch <-  fish_catch[-1, ]   
fish_catch$year <- as.numeric(fish_catch$year)
fish_catch$wild <- as.numeric(fish_catch$wild)

fish_catch <- fish_catch %>% 
  mutate(Year_from_1950 = year-1950)

ggplot(fish_catch, aes(x = year)) +
  geom_point(aes(y = wild)) +
  theme_classic()


```

The data looks like it might fit a logistic growth model fairly well. Catch seems to increase exponentially initially in the series and then level off to a value around 90 million tons

$N(t) = \frac{A}{(1+Be^{-rt})}$
$B = \frac{(K-N0)}{N0}$

Estimate of K = 90
Estimate of N0 = 17
Estimate of B = 4.3


exponential before 1970, or time = 20

```{r message=FALSE, warning=FALSE, fig.align='center'}
exponential <- fish_catch %>% 
  filter(Year_from_1950<20)

ggplot(exponential, aes(x = year, y = log(wild))) +
  geom_point()

lm(log(wild) ~ year , data = fish_catch) #r ~ 0.025

```

 r ~ 0.025


```{r}


fish_fit <- nls(wild ~ A/(1+B*exp(-r*Year_from_1950)), 
                start = list(A = 90, B = 4.3, r = 0.025), 
                data = fish_catch, trace = FALSE) 


summary(fish_fit)


A <- coef(fish_fit)[1]
B <- coef(fish_fit)[2]
r <- coef(fish_fit)[3]


```

```{r fig.align='center'}


fish_pred <- A/(1+ B*exp(-r*fish_catch$Year_from_1950))

pred_df <-  data.frame(year = fish_catch$year, fish_pred)


ggplot(fish_catch, aes(x = year, y = wild)) +
  geom_point( color = "steelblue") +
  geom_line(data = pred_df, aes(x = year, y = fish_pred), color = "mediumaquamarine") +
  labs(x = "Year", y = "Wild Catch (Million Tons)") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5)) +
  ggtitle("Global Wild Fish Catch 1950 - 2010 with Parameterized Model")

```


### Task 3: Bootstrapped Confidence Interval for Proportions


```{r fig.align='center'}

nonbinary <- data.frame(n = seq(1,36), hostile = c(rep(1,22),rep(0,36-22)))


test_non <- nonbinary %>% 
  group_by(hostile) %>% 
  count() %>% 
  mutate(prop = nn/length(nonbinary$hostile))


proportion <- function(x,i){
  
    sum(x[i] == 1)/length(x[i])
  
}

hostile_boot <- boot(nonbinary$hostile, statistic = proportion, R=100000)
hostile_boot_results <- data.frame(proportion = hostile_boot$t)
hostile_boot


ggplot(hostile_boot_results, aes(x = proportion)) +
  geom_histogram(binwidth = 0.05, fill = "slategrey") +
  theme_classic()


mean_proportion <- mean(hostile_boot_results$proportion)
high_end <- mean_proportion + 1.96*(sd(hostile_boot$t))
low_end <- mean_proportion - 1.96*(sd(hostile_boot$t))

mean_proportion
high_end
low_end

```


Data from 2014 UCSB Campus Climate Project Final Report (prepared by
Rankin & Associates Consulting) show that 22 out of 36 surveyed UCSB community members (61%) identifying as nonbinary/genderqueer responded that they had personally experienced “exclusionary, offensive, hostile or intimidating conduct” (compared to 26% and 19% for those identifying as women and men, respectively).

Using this reponse data (n=36) from nonbinary/genderqueer respondants, a 95% confidence interval was calculated by bootstrapping 100,000 times the proportion of repondants who respond yes. The 95% confidence interval ranges from 45.1% to 77.1%



### Task 4: Watch 3 RStudio::conf talks


Talk 1: Visualizing uncertainty with hypothetical outcomes plots

Speaker: Claus Wilke

Notes:

- Learned how to generate animated graphics for uncertain analyses(Hypothetical Outcome Plots, HOPS). Different lines draws for each "draw" of an uncertainty analysis.

- Psychologically it seems that no transition between different "runs" of an uncertainty analysis is easier to comprehend

Question:

In what circumstances do you think that showing an animated Hypothetical Outcome Plot is inappropriate for an audience?

Talk 2: Melt the clock Tidy time series analysis 

Speaker: Earo Wang

Notes:

- 2 different packages for working with time series in a tidy format (tsibble and fable). These packages are made to deal with the weird model formats of time series data. Looks super useful for making easy time series, I want to use it on my GP/shiny app data!

-appears that you can still do the same forecasting analyses with her fable package, makes the syntax for models really easy!

Question:

Can you forecast data using these packages with data that incorporate multiple seasonalities? 

Talk 3: Cultivating creativity in data work

Speaker: Hilary Parker

Notes:

Question:
